Wigglyhedra

### 2024-07-14 draft/wigglyhedra.tex
Edited categorical section.

\begin{proposition}
\label{prop:ext1}
  Consider two wiggly arcs \(\alpha\) and \(\alpha'\), and let \(X\) and \(X'\) be objects in the standard heart \(\mathcal{A}_m\) corresponding to \(\alpha\) and \(\alpha'\) respectively.
  Then \(\alpha\) and \(\alpha'\) are compatible in the sense of~\cref{def:compatible} if and only if \(\Ext^1(X, X') = 0\).
  More generally, the incompatibility degree \(\delta(\alpha,\alpha')\) as in~\cref{def:incompatibilityDegree} is the dimension of \(\Ext^1(X,X')\).
\end{proposition}

\begin{proof}
  The proof follows from a specialisation of the bigraded intersection number calculations that appear in Sections 3 and 4 of~\cite{kho.sei:02}.
  Let \(X\) and \(X'\) be objects in the standard heart that correspond to wiggly arcs \(\alpha\) and \(\alpha'\).
  The bigraded intersection number as defined in~\cite[Sect.~3d]{kho.sei:02} between the curves \(\alpha\) and \(\alpha'\) is a polynomial in the variables \(q_1^{\pm}\), \(q_2^{\pm}\).
  It follows from~\cite[Prop.~4.9]{kho.sei:02} that this bigraded intersection number equals the Poincar{\'e} polynomial of the morphisms between \(X\) and \(X'\), namely the expression
  \begin{equation}
    \label{eq:bigraded-homs}
  \sum_{s_1,s_2}\dim(\Hom_{\mathcal{C}_m}(X,X'\langle -s_2 \rangle[s_1]))q_1^{s_1}q_2^{s_2}.    
\end{equation}
  Note that a morphism in \(\Hom_{\mathcal{C}_m}(X,X'\langle -s_2 \rangle[s_1])\) has level \(s_1 - s_2\); that is, it takes the object \(X\) in level \(0\) to the object \(X'\langle -s_2 \rangle[s_1]\) in level \(s_1 - s_2\).
  We can obtain the dimension of all morphisms of level \(\ell\) by setting \(q_1 = q\) and \(q_2 = q^{-1}\) and taking the coefficient of \(q^{\ell}\) in the expression~\eqref{eq:bigraded-homs}.
    The dimension of \(\Ext^1(X,X')\), is therefore the coefficient of \(q\) under this specialisation.

However, the grading convention in~\cite{kho.sei:02} is not the same as ours --- our grading is by path length, while the grading in~\cite{kho.sei:02} is one for arrows in one direction and zero for arrows in the opposite direction.
So we have to modify the gradings when we construct the complexes, and thus the morphisms that now have level one change as well.
In particular, the table in~\cite[Lem.~3.20]{kho.sei:02} that computes the bigraded intersection numbers will change, and so will the table in~\cite[Lem.~4.12]{kho.sei:02} that computes the Poincar{\'e} polynomials of \(\Hom(P_k,-)\).

It can be checked that under the modified grading, the space of morphisms arising from an intersection of curves of ``crossing'' type as well as ``non pointed'' type contains exactly one morphism (up to equivalence) of level one; that is, a unique element of \(\Ext^1\).
Morphisms arising from other intersections do not contribute to \(\Ext^1\).
This calculation proves the proposition.
\end{proof}

>>>>

Recall that two wiggly arcs \(\alpha\) and \(\alpha'\) are said to be compatible (as in~\cref{def:compatible}) if they are non-crossing and pointed.
The incompatibility degree of \(\alpha\) and \(\alpha'\) (as in~\cref{def:incompatibilityDegree}) is \(1\) if \(\alpha\) and \(\alpha'\) are non pointed, and the number of crossings of \(\alpha\) and \(\alpha'\) otherwise.
\begin{proposition}
\label{prop:ext1}
Let \(X\) and \(X'\) be objects in the standard heart \(\mathcal{A}_m\) corresponding to two wiggly arcs \(\alpha\) and \(\alpha'\) respectively.
  Then \(\alpha\) and \(\alpha'\) are compatible if and only if \(\Ext^1(X, X') = 0\).
  More generally, the incompatibility degree of \(\alpha\) and \(\alpha'\) is the dimension of \(\Ext^1(X,X')\).
\end{proposition}

\begin{proof}
  The proof follows from a specialisation of the bigraded intersection number calculations that appear in Sections 3 and 4 of~\cite{kho.sei:02}.
  The bigraded intersection number as defined in~\cite[Sect.~3d]{kho.sei:02} between the curves \(\alpha\) and \(\alpha'\) is a polynomial in the variables \(q_1^{\pm}\), \(q_2^{\pm}\).
  It follows from~\cite[Prop.~4.9]{kho.sei:02} that this bigraded intersection number equals the Poincar{\'e} polynomial of the morphisms between \(X\) and \(X'\), namely the expression
  \begin{equation}
    \label{eq:bigraded-homs}
  \sum_{s_1,s_2}\dim(\Hom_{\mathcal{C}_m}(X,X'\langle -s_2 \rangle[s_1]))q_1^{s_1}q_2^{s_2}.    
\end{equation}
However, the grading convention in~\cite{kho.sei:02} is not the same as ours --- our grading is by path length, while the grading in~\cite{kho.sei:02} is one for arrows in one direction and zero for arrows in the opposite direction.
So we have to modify the gradings when we construct the complexes.
In particular, the table in~\cite[Lem.~3.20]{kho.sei:02} that computes the bigraded intersection numbers will change, and so will the table in~\cite[Lem.~4.12]{kho.sei:02} that computes the Poincar{\'e} polynomials of \(\Hom(P_k,-)\).

With our path-length grading, note that a morphism in \(\Hom_{\mathcal{C}_m}(X,X'\langle -s_2 \rangle[s_1])\) has level \(s_1 - s_2\); that is, it takes an object \(X\) of level \(0\) to the object \(X'\langle -s_2 \rangle[s_1]\) of level \(s_1 - s_2\).
  We can thus obtain the dimension of all morphisms of level \(\ell\) by setting \(q_1 = q\) and \(q_2 = q^{-1}\) and taking the coefficient of \(q^{\ell}\) in the expression~\eqref{eq:bigraded-homs}.
    The dimension of \(\Ext^1(X,X')\), is therefore the coefficient of \(q\) under this specialisation.

It can be checked that under the path-length grading, the space of morphisms arising from an intersection of curves of ``crossing'' type as well as ``non pointed'' type contains exactly one morphism (up to equivalence) of level one; that is, a unique element of \(\Ext^1\).
Morphisms arising from other intersections do not contribute to \(\Ext^1\).
This calculation proves the proposition.
\end{proof}
